**The status of the microbial census: an update**

### 2015-12-04 Schloss_Census2_mBio_2015.Rmd
Add phylum by category supplemental tables

**Figure results/figures/time_course_figure.pdf. Progression of the microbial census since the first full-length 16S rRNA gene sequence was deposited into GenBank in 1983.*** The number of bacterial and archaeal 16S rRNA gene sequences deposited (A) and the new OTUs they represent (B) has increased exponentially until the last several years when the rate of change has plateaued. For both bacterial and archaeal sequences, the number of studies that are responsible for depositing more than 50% of the seuqences each year has been relatively small (C).

**Figure results/figures/category_phylum_heatmap.pdf. Heatmap depicting the relative abundance of the most common bacterial and archaeal phyla across different environments.** Each environmental category exhibited a phylum-level signature although the bacterial census was dominated by sequences from the Firmicutes, Proteobacteria, Actinobacteria, and Bacteroidetes and the archaeal census was dominated by sequences from the Euryarchaeota and Thaumarchaeota.

@@

```{r figure4a, eval=FALSE}
data <- read.table(file="data/mothur/all_bacteria.env_category.groups.rarefaction", header=T)
category <- "\\.plant$" #change the SO to whatever category you want (leave the \\. and $)
plant_columns <- grep(category, colnames(data))
category <- "\\.aerosol$" #change the SO to whatever category you want (leave the \\. and $)
aerosol_columns <- grep(category, colnames(data))
category <- "\\.aquatic$" #change the SO to whatever category you want (leave the \\. and $)
aquatic_columns <- grep(category, colnames(data))
category <- "\\.built$" #change the SO to whatever category you want (leave the \\. and $)
built_columns <- grep(category, colnames(data))
category <- "\\.other$" #change the SO to whatever category you want (leave the \\. and $)
other_columns <- grep(category, colnames(data))
category <- "\\.soil$" #change the SO to whatever category you want (leave the \\. and $)
soil_columns <- grep(category, colnames(data))
category <- "\\.zoonotic$" #change the SO to whatever category you want (leave the \\. and $)
zoonotic_columns <- grep(category, colnames(data))
plant_aerosol_aquatic_built_other_soil_zoonotic_data <- data[,c(1, plant_columns, aerosol_columns, aquatic_columns, built_columns, other_columns, soil_columns, zoonotic_columns)]  #the 1 gets the first column which is numsampled
nrows_to_keep <- which.max(plant_aerosol_aquatic_built_other_soil_zoonotic_data$numsampled[!is.na(plant_aerosol_aquatic_built_other_soil_zoonotic_data[,2])])
plant_aerosol_aquatic_built_other_soil_zoonotic_data <- plant_aerosol_aquatic_built_other_soil_zoonotic_data[1:nrows_to_keep,]
pdf(file="/Users/renegirard7gfy/Desktop/Bac_Uniq_3per.pdf")

par(mar=c(5,5,0.5,0.5))
plot(NA, type="n", ylim=c(0,max(40000)),
     xlim=c(0,max(data$numsampled)),
     xlab="Number of sequences sampled",
     ylab="",
     axes=FALSE)

points(data$X0.03.plant~data$numsampled, type="l", col="red", lwd=2)
points(data$X0.03.aerosol~data$numsampled, type="l", col="green", lwd=2)
points(data$X0.03.aquatic~data$numsampled, type="l", col="blue", lwd=2)
points(data$X0.03.built~data$numsampled, type="l", col="orange", lwd=2)
points(data$X0.03.other~data$numsampled, type="l", col="black", lwd=2)
points(data$X0.03.soil~data$numsampled, type="l", col="purple", lwd=2)
points(data$X0.03.zoonotic~data$numsampled, type="l", col="deeppink", lwd=2)

legend(x=5e5, y=20000, lwd=2, lty=1,
       legend=c("Other", "Plant", "Aerosol", "Aquatic", "Built", "Soil", "Zoonotic"),
       col=c("black", "red", "green", "blue", "orange", "purple", "deeppink"))

axis(1, at=seq(0,700000,1e5), label=c("0", "100,000", "200,000", "300,000", "400,000", "500,000", "600,000", "700,000"), cex.axis=0.8)
axis(2, at=seq(0,40000,5000), label=c("0", "5,000", "10,000", "15,000", "20,000", "25,000", "30,000", "35,000", "40,000"), las=2, cex.axis=0.8)
box()
mtext(side=2, "Number of OTUs", line=4)

dev.off()

**Figure 4B. Archaeal sampling by environment**

```{r figure4b, eval=FALSE}

data <- read.table(file="data/mothur/all_archaea.env_category.groups.rarefaction", header=T)
category <- "\\.plant$" #change the SO to whatever category you want (leave the \\. and $)
plant_columns <- grep(category, colnames(data))
category <- "\\.aerosol$" #change the SO to whatever category you want (leave the \\. and $)
aerosol_columns <- grep(category, colnames(data))
category <- "\\.aquatic$" #change the SO to whatever category you want (leave the \\. and $)
aquatic_columns <- grep(category, colnames(data))
category <- "\\.built$" #change the SO to whatever category you want (leave the \\. and $)
built_columns <- grep(category, colnames(data))
category <- "\\.other$" #change the SO to whatever category you want (leave the \\. and $)
other_columns <- grep(category, colnames(data))
category <- "\\.soil$" #change the SO to whatever category you want (leave the \\. and $)
soil_columns <- grep(category, colnames(data))
category <- "\\.zoonotic$" #change the SO to whatever category you want (leave the \\. and $)
zoonotic_columns <- grep(category, colnames(data))
plant_aerosol_aquatic_built_other_soil_zoonotic_data <- data[,c(1, plant_columns, aerosol_columns, aquatic_columns, built_columns, other_columns, soil_columns, zoonotic_columns)]  #the 1 gets the first column which is numsampled
nrows_to_keep <- which.max(plant_aerosol_aquatic_built_other_soil_zoonotic_data$numsampled[!is.na(plant_aerosol_aquatic_built_other_soil_zoonotic_data[,2])])
plant_aerosol_aquatic_built_other_soil_zoonotic_data <- plant_aerosol_aquatic_built_other_soil_zoonotic_data[1:nrows_to_keep,]
pdf(file="/Users/renegirard7gfy/Desktop/Arc_Uniq_3per.pdf")

par(mar=c(5,5,0.5,0.5))
plot(NA, type="n", ylim=c(0,max(data[,-1], na.rm=T)),
     xlim=c(0,max(data$numsampled)),
     xlab="Number of sequences sampled",
     ylab="",
     axes=FALSE)

points(data$X0.03.plant~data$numsampled, type="l", col="red", lwd=2)
points(data$X0.03.aerosol~data$numsampled, type="l", col="green", lwd=2)
points(data$X0.03.aquatic~data$numsampled, type="l", col="blue", lwd=2)
points(data$X0.03.built~data$numsampled, type="l", col="orange", lwd=2)
points(data$X0.03.other~data$numsampled, type="l", col="black", lwd=2)
points(data$X0.03.soil~data$numsampled, type="l", col="purple", lwd=2)
points(data$X0.03.zoonotic~data$numsampled, type="l", col="deeppink", lwd=2)

legend(x=15000, y=1550, lwd=2, lty=1,
       legend=c("Other", "Plant", "Aerosol", "Aquatic", "Built", "Soil", "Zoonotic"),
       col=c("black", "red", "green", "blue", "orange", "purple", "deeppink"))

axis(1, at=seq(0,35000,5000), label=c("0", "5,000", "10,000", "15,000", "20,000", "25,000", "30,000", "35,000"), cex.axis=0.8)
axis(2, at=seq(0,5250,750), label=c("0", "750", "1500", "2250", "3,000", "3750", "4500", "5250"), las=2, cex.axis=0.8)
box()
mtext(side=2, "Number of OTUs", line=4)

dev.off()
```

>>>>

**Figure results/figures/time_course_figure.pdf. Progression of the microbial census since the first full-length 16S rRNA gene sequence was deposited into GenBank in 1983.*** The number of bacterial and archaeal 16S rRNA gene sequences deposited (A) and the new OTUs they represent (B) has increased exponentially until the last several years when the rate of change has plateaued. For both bacterial and archaeal sequences, the number of studies that are responsible for depositing more than 50% of the sequences each year has been relatively small (C).

**Figure results/figures/category_phylum_heatmap.pdf. Heatmap depicting the relative abundance of the most common bacterial and archaeal phyla across different environments.** Each environmental category exhibited a phylum-level signature although the bacterial census was dominated by sequences from the Firmicutes, Proteobacteria, Actinobacteria, and Bacteroidetes and the archaeal census was dominated by sequences from the Euryarchaeota and Thaumarchaeota. The ten most abundant phyla across all environmental categories are shown. The data for all bacterial and archaeal phyla are available in Supplemental Tables 2 and 3, respectively.

@@

### 2015-12-04 results/tables/build_bacterial_category_phylum_table.Rmd
Add phylum by category supplemental tables

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
tables: true
geometry: margin=1.0in
classoption: landscape
---

**Supplementary Table 3. Frequency that each archaeal phylum was found across each of the environmental categories.**

@@

archaea <- data[data$domain=="archaea",-1]
archaea <- archaea[order(archaea$total, decreasing=T), names(categories)]
colnames(archaea) <- gsub("total", "Total", colnames(archaea))
rownames(archaea) <- gsub("Candidate_division_", "", rownames(archaea))
rownames(archaea) <- gsub("_", " ", rownames(archaea))

total <- apply(archaea, 2, sum)
archaea <- rbind(archaea, Total=total)
archaea <- cbind(rownames(archaea), archaea)

colnames(archaea) <- c("", categories)
composite <- xtable(archaea)

align(composite) <- paste0('ll', paste(rep('c', ncol(archaea)-1), collapse=""))

>>>>

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
tables: true
geometry: margin=0.1in
classoption: landscape
---

**Supplementary Table 2. Frequency that each bacterial phylum was found across each of the environmental categories.**

@@

bacteria <- data[data$domain=="bacteria",-1]
bacteria <- bacteria[order(bacteria$total, decreasing=T), names(categories)]
colnames(bacteria) <- gsub("total", "Total", colnames(bacteria))
rownames(bacteria) <- gsub("Candidate_division_", "", rownames(bacteria))
rownames(bacteria) <- gsub("_", " ", rownames(bacteria))

total <- apply(bacteria, 2, sum)
bacteria <- rbind(bacteria, Total=total)
bacteria <- cbind(rownames(bacteria), bacteria)

colnames(bacteria) <- c("", categories)
composite <- xtable(bacteria)

align(composite) <- paste0('ll', paste(rep('c', ncol(bacteria)-1), collapse=""))
